Lesson 6
========================================================

### Welcome

***

### Scatterplot Review

```{r Scatterplot Review}
library('ggplot2')
data(diamonds)
qplot(data=diamonds,x=carat,y=price,xlim=c(0,quantile(diamonds$carat,0.99)),ylim=c(0,quantile(diamonds$price,0.99)))+geom_point(fill=I('#F79420'),color=I('black'),shape=21)+stat_smooth(method='lm')

```

***

### Price and Carat Relationship
Response: Dispersion of the prices increases as carat value increases. This may be linear or exponential. We can judge by the smoothening function that it is not completely linear.

### ggpairs Function
Notes: Sampling the data is better for plotting since it's a huge dataset.

```{r ggpairs Function}
# install these if necessary
install.packages('GGally')
install.packages('scales')
install.packages('memisc')
install.packages('lattice')
install.packages('MASS')
install.packages('car')
install.packages('reshape')
install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

What are some things you notice in the ggpairs output?
Response: The previous scatter plot and newly generated box-plots indicate the relation between carat and price. And since carat/volume is a cube value(xyz), taking its cube-root converts it into a linear function. 

***

### The Demand of Diamonds
Notes:

```{r The Demand of Diamonds}
library(gridExtra)
plot1 <- qplot(data=diamonds,x=price,binwdith=100,fill=I('#099009'))+ggtitle('Price')

plot2 <- qplot(data=diamonds,x=price,binwdith=0.01,fill=I('#F79420'))+scale_x_log10()+ggtitle('Price (log10)')

grid.arrange(plot1,plot2,ncol=2)
```

***

### Connecting Demand and Price Distributions
Notes:log10 removes the skewed nature of the data,therefore it indicates that count depended on the price more accurately when deicded with many other factors such as inflation,time,rich-to-poor ratio,etc

***

### Scatterplot Transformation

```{r Scatterplot Transformation}
qplot(data=diamonds,x=carat,y=price)+scale_y_log10()+geom_point()
```


### Create a new function to transform the carat variable

```{r Cuberoot Transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat),decreasing=T))
head(sort(table(diamonds$price),decreasing=T))

```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(position='jitter',alpha=1/2,size=3/4) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:After seeing how initially, the plot was overcrowded, and by checking the density by Sorting and Heading the table, we add jitter, alpha and size to th geon_point to fix Overplotting.
Clarity also increases price.
Cut also acts as a factor to price.

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
install.packages('RColorBrewer',dependencies=TRUE)
library(RColorBrewer)

ggplot(aes(x = carat, y = price,color=clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response: As clarity increases, price increases. 

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response: There's not much variation of price wrt cut,
Despite the fact that cut influences the diamond's sparkle.

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title ='Color', reverse = FALSE,override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response: Color affects Pirce just like Clarity. Color differences are not that noticable, but still make a difference.

***

### Linear Models in R
Response: We use linear regression here, since the output variable, Price, is of continuous values, and is affected by more than one input variable.

***

### Building the Linear Model
Notes: We choose are ouput as log(price) , since we are predicting diamonds prices, and we select our input/explanatory variable as cube-root of carat, since it has previously shown a linear relation with price.


```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits=3)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Problems with model
Response:  
1) data set is old, therefore current inflation not considered. 
2) diamond prices have changed due to different new factors within carat itself, therefore the old data might not support predictions for modern-day prices.

***

### A Bigger, Better Data Set
Notes: Import and use the latest diamond dataset.

```{r A Bigger, Better Data Set}
install.packages('bitops')
install.packages('RCurl')
library('bitops')
library('RCurl')
# diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
# load(rawConnection(diamondsurl))
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

```{r Building a Model Using the Big Diamonds Data Set}
diamondsBig <- read.csv('diamondsbig.csv')
diamondsBig$logprice=log(diamondsBig$price)
m1 <- lm(I(logprice) ~ I(carat^(1/3)),
         data=diamondsBig[diamondsBig$price<10000 & diamondsBig$cert=="GIA",])
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits=3)

```


***

## Predictions
We try an example:
Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
library(memisc)
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
exp(modelEstimate)
```
***
## Final Thoughts
Notes: The model predicts a price slightly higher than the actual value but well within the expected price.

Actual Price: 5000
Predicted Price: 5040

This is within the 95% confidence interval, we used to predict the price.
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

